Sequence Specific Modeling of \emph{E.~coli} Cell-Free Protein Synthesis

### 2017-04-07 Paper_v1.tex
Working my way thru the results .... still ...

\subsection{Model derivation and validation}
The cell-free stoichiometric network was constructed by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa},
and adding deletions associated with the specific cell free system (see materials and methods).
We then added the transcription and translation template reactions of Allen and Palsson for the specific proteins of interest \cite{Allen:2003aa}.
A schematic of the metabolic network, which consisted of 264 reactions and 146 species, is shown in Fig.~\ref{fig:network}A.
The network described the major carbon and energy pathways, as well as amino acid biosynthesis and degradation pathways.
Using this network, in combination with effective promoter models taken from Moon et al. [REF], and literature values for cell-free culture parameters (Table ~\ref{tbl:parameters}),
we simulated the sequence specific production of two model proteins, chloramphenicol acetyltransferase (CAT) and dual emission green fluorescent protein (deGFP) using
different \textit{E.~coli} cell-free extracts.
The cell free metabolic network, along with all model code and parameters, can be downloaded under an MIT software license from the Varnerlab GitHub site [REF].

@@

\subsubsection{Productivity}
The theoretical maximum mean productivity ($\mu$M/h) was inversely proportional to the protein length ($l_{P}$) and varied between 1 and 12 $\mu$M/h for the proteins sampled (Fig.~\ref{fig:Prof}A-B). The theoretical maximum productivity was within a standard deviation with and without
amino acid supplementation for each protein, but varied significantly between proteins.
For instance, BMP10 (XX aa) had a optimal productivity of approximately 2.5 $\mu$M/h whereas CAT (ZZ aa) had an optimal productivity of approximately 12 $\mu$M/h.
To examine the influence of protein length further, the mean optimal productivity was plotted against the carbon number of each protein (Fig.~\ref{fig:Prof}B).
The optimal productivity and protein length were related by the power-law relationship $\alpha\times(l_{P})^{\beta}$, where $\alpha = 6.0\times 10^{6}$ [UNITS?] and $\beta=-1.9$.
Interestingly, CAT did not obey the power-law relationship; the relatively high productivity of CAT was due to its T7 promoter.
However, CAT expressed under a P70a promoter followed the power-law correlation with a productivity of approximately 8.8 $\mu$M/h (predicted to be XX by the optimal productivity correlation). Taken together, the maximum optimal productivity of a cell free reaction was found to be inversely proportional to protein size,
following a power-law relationship for proteins expressed under a P70a promoter.

@@

\subsubsection{Carbon Yield}
The theoretical maximum mean carbon yield was inversely proportional to protein length and varied between 40\% and 57\% for the proteins sampled (Fig.~\ref{fig:Prof}E-F).
The relationship between the optimal carbon yield and protein length was linear; $m_{Y}\times(l_{P})+b_{Y}$ where $m_{Y} =  -8.03\times10^{-3}$ [UNITS] with supplementation, and $m_{Y} = -6.98\times10^{-3}$ [UNITS] without. The y-intercept $b_{Y} = 63.83$ [UNITS] with supplementation, and $b_{Y} = 58.86$ [UNITS] without.
The linear carbon yield models showed good predictability for estimating the optimal carbon yield for a range of protein lengths ($R^{2} = 0.95$ with amino acids
and $R^{2} = 0.90$ without supplementation), irrespective of the promoter used to control protein expression.
The effective yield models were approximately parallel, with a drop in carbon yield of approximately 7\% without supplementation.
Only the necessary amount of amino acids were used for the production of the protein of interest; thus, it may be hypothesized that all the glucose was used to power CFPS and did not contribute to the carbon yield.
In that case, the carbon yield without the glucose contribution would be 100\%.
However, in the experimentally constrained case, CAT was produced with a carbon yield of 6.2\% compared to the theoretical maximum of 58\%.
This decrease in carbon yield suggests inefficiencies in CFPS that can potentially be improved.
ssFBA assumes a psuedo steady state; thus, intermediate metabolites cannot accumulate within the cell-free extract.
In addition, ssFBA is solved by maximizing the flux through the protein production reaction.
Therefore, carbon flux will travel through the network to optimize the maximum flux through the protein synthesis reaction.
In examining the experimental dataset, there is a high accumulation of organic acids, especially acetate, pyruvate and lactate.
The experimental performance could be improved by diverting this carbon toward the protein of interest by knockouts during the cell-free extract preparation.

@@

To investigate the connection between carbon yield and oxidative phosphorylation further,
we calculated the CAT carbon yield as a function of the oxidative phosphorylation flux (Fig.~\ref{fig:oxphos_yield}).
Oxidative phosphorylation had a strong effect on the carbon yield, both with and without amino acid supplementation.
In the presence of amino acid supplementation, the carbon yield ranged from 20\% to approximately 60\%,
depending upon the oxidative phosphorylation flux.
However, without amino acid supplementation, the carbon yield dropped to approximately 10\%, and reached a maximum of 50\%.
In the absence of supplementation, a lower carbon yield was expected for the same oxidative phosphorylation flux, as glucose was utilized for both energy generation and amino acid biosynthesis.
In all cases, whenever the carbon yield was below its theoretical maximum, there was an accumulation of both acetate and lactate.
The experimental dataset also exhibited a mixture of acetate and lactate accumulation during CAT synthesis, which suggested the CFPS reaction was not
operating with optimal oxidative phosphorylation activity.
Oxidative phosphorylation is a membrane associated process, however CFPS has no cell membrane.
Jewett et al hypothesized that membrane vesicles present in the CFPS reaction carry out oxidative phosphorylation [REF];
however, the number, size, protein loading and lifetime of these vesicles remains an open area of study.
Thus, we expect CFPS will have suboptimal oxidative phosphorylation capacity.

% The addition of phosphate showed an increase in CAT yield; however, it is unclear whether the addition of phosphate enhances oxidative phosphorylation, inhibits phosphatase reactions, or both \cite{Jewett:2008aa}.
% Interestingly, the addition of NADH did not increase the rate of protein synthesis, since the concentration of ATP was most likely saturated.
% An alternative strategy may be the inhibition of anaerobic processes in cell-free, in order to minimize unwanted byproducts such as acetate and lactate.
%Thus, it is interesting to see how much of the energetic needs of the system are met by oxidative phosphorylation and how much from anaerobic processes.
\begin{figure}[t!]
\includegraphics[width=0.7\textwidth]{./figs/Fig-4-OxyPhox-versus-Carbon-Yield.pdf}
\caption{deGFP carbon yield versus oxidative phosphorylation flux, across an ensemble of 1000 ssFBA solutions, for three cases: amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray).}
\label{fig:oxphos_yield}
\end{figure}

\subsection{Optimal Metabolic Flux Distribution}
Amino acid supplementation altered the optimal metabolic flux distribution predicted for CAT production (Fig.~\ref{fig:flux}).
To investigate the influence of amino acid supplementation, we compared the simulated metabolic flux distributions for CAT production with and without external amino acids.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-5-FluxDistribition-Optimal.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and oxidative phosphorylation, for three different cases: (A) amino acid uptake and synthesis, (B) amino acid uptake without synthesis, and (C) amino acid synthesis without uptake. Mean flux across the ensemble, normalized to glucose uptake flux. Thick arrows indicate flux to or from amino acids.}
\label{fig:flux}
\end{figure}
In the presence of amino acid supplementation, and \textit{de~novo} amino acid synthesis, there was an incomplete TCA cycle, where a combination
of glucose and amino acids powered protein expression (Fig.~\ref{fig:flux}A).
Glucose was consumed to produce acetyl-coenzyme A, and associated by-products,
while glutamate was converted to alpha-ketoglutarate which traveled to oxaloacetic acid and pyruvate for additional amino acid biosynthesis.
On the other hand, in the presence of amino acid supplementation, but without \textit{de~novo} amino acid biosynthesis, there was no TCA cycle flux.
In this case, ATP was produced by a combination of substrate level and oxidative phosphorylation, where
ubiquinone was regenerated via \textit{nuo} activity, without relying on succinate dehydrogenase in the TCA cycle (Fig.~\ref{fig:flux}B).
These first two cases where amino acids were available had similar performance, and their respective metabolic flux distributions had a 99\% correlation for all proteins. However, in the absence of amino acid supplementation (where all amino acids were synthesized \textit{de~novo} from glucose),
the energy efficiency and carbon yield decreased; in this case the TCA cycle was largely complete and there was diversion of metabolic flux into the Entner-Doudoroff pathway to produce NADPH (Fig.~\ref{fig:flux}C). [Mike: we need a link sentence to the experimentally constrained case.]

@@

The case constrained by experimental measurements (Fig. \ref{fig:flux_exp}), had the highest correlation of 0.66 with the flux distribution from the case supplied with no amino acids and a correlation of 0.52 with the cases supplied with amino acids.
Thus there are some differences in the flux distribution compared to the optimum solutions which may provide some insight to improve CFPS performance.
Metabolic fluxes were constrained by experimental measurements (available in Supporting Information) where available for the first hour which constrained the solution space of ssFBA to have a more realistic depiction of the flux distribution.
The central carbon organic acids showed good agreement with the data (Fig.~\ref{fig:flux_exp}B).
Only certain amino acid synthesis reactions were blocked since during the growth of \emph{E.~coli} not all amino acids were supplied (see Materials and Methods).
During the cell-free reaction all amino acids were supplied, however glucose still traveled through all the major pathways, and the same metabolic precursors were still utilized for amino acid biosynthesis.
In this case, it is unclear which substrate (glucose or amino acids) is used to power CFPS and may in fact be a combination of both.
The optimum solutions only produced the required amount of amino acids necessary, however in examining the measurements there is an accumulation of alanine and glutamine which may explain some of the differences in the correlations of the flux distributions.
Accumulation of pyruvate, lactate, acetate, and other organic acids can be seen (Fig.~\ref{fig:flux_exp}B) and shows an inefficiency of carbon utilization.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-6-FluxDistribition-Experimental.pdf}
\caption{ssFBA simulation of CAT production for an experimentally constrained case. (A) Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and oxidative phosphorylation.  Mean flux across the ensemble, normalized to glucose uptake flux. Thick arrows indicate flux to or from amino acids. (B) Central carbon metabolite measurements versus ssFBA simulations over a one hour time course.}
\label{fig:flux_exp}
\end{figure}

\subsubsection{Potential Alternative Metabolic Optima}
One of the interesting features of constraint based approaches is the ability to compute potentially alternative optimal solutions.
These solutions have the same objective value, e.g., the productivity but different underlying metabolic flux distributions.
Despite the constraints by experimental measurements, it is difficult to calculate the physiological flux distribution of metabolism (Fig.~\ref{fig:norm})
For example, there is a high flux through the Entner-Douodoroff pathway, but this is likely non-physiological, and simply an artifact of the optimal solution of ssFBA.
A knockout of the Entner-Doudoroff pathway has no effect on the norm productivity of CAT compared to no knockouts (Fig.~\ref{fig:norm}A).
In addition, pairwise knockouts of Entner-Doudoroff and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) in the network result in the same optimal solution of CAT productivity.
However, there is a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); ssFBA will reroute the flux to optimize the objective function.
Interestingly, there are crucial subgroups that are detrimental to CAT productivity.
For example, a single group knockout of glycolysis/gluconoeogenesis or oxidative phosphorylation leads to no CAT production.
There were also cases of group knockouts that had no effect on productivity or its flux distribution such as isoleucine, leucine, and valine biosynthesis knocked out with histidine biosynthesis.
Flux balance analysis has been shown to have multiple alternative optimal solutions with flux variablitiy analysis and mixed-integer linear programming resulting in a poor depiction of the physiological flux distribution \cite{LEE2000711, Mahadevan2003264, Schuetz119}.
In our case, ssFBA reached the same optimal solution of CAT productivity for 39.9\% of the pairwise group knockouts, but had a different flux distribution than compared to no knockouts.
To determine which reactions occur in CFPS, adding thermodynamic feasibility constraints to reactions may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013}.
It would also be interesting to track the carbon flux using C\textsuperscript{13} labeling in CFPS and constrain branch reactions in ssFBA to the resulting measurements, a method that has been shown to represent the flux distribution for \emph{in~vivo} processes well \cite{Zamboni:2009}.
\begin{figure}[t!]
\includegraphics[width=1.0\textwidth]{./figs/Fig-7-FluxDistribition-Analysis.pdf}
\caption{The norm effect of pairwise knockouts of subgroups in the cell-free network (A) CAT productivity norm compared to no knockouts in the experimentally constrained case. (B) Flux distribution norm compared to no knockouts in the experimentally constrained case.}
\label{fig:norm}
\end{figure}

@@

\subsection*{Pairwise group knockouts.}
Pairwise and single group knouckouts were simulated in ssFBA by setting the flux bounds for the all the reactions in a group to zero.
We grouped reactions in the cell-free network into 19 subgroups (available in Supporting Information).
We computed the norm of the productivity of CAT for each pairwise knockout compared to the productivity of CAT with no knockouts.
We also computed the norm of the flux distribution for each pairwise knockout compared to the flux distribution with no knockouts.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

>>>>

\subsection{Model derivation and validation}
The cell-free stoichiometric network was constructed by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa},
and adding deletions associated with the specific cell free system (see materials and methods).
We then added the transcription and translation template reactions of Allen and Palsson for the specific proteins of interest \cite{Allen:2003aa}.
A schematic of the metabolic network, which consisted of 264 reactions and 146 species, is shown in Fig.~\ref{fig:network}A.
The network described the major carbon and energy pathways, as well as amino acid biosynthesis and degradation pathways.
Using this network, in combination with effective promoter models taken from Moon et al. [TODO:REF], and literature values for cell-free culture parameters (Table ~\ref{tbl:parameters}),
we simulated the sequence specific production of two model proteins, chloramphenicol acetyltransferase (CAT) and dual emission green fluorescent protein (deGFP) using
different \textit{E.~coli} cell-free extracts.
The cell free metabolic network, along with all model code and parameters, can be downloaded under an MIT software license from the Varnerlab GitHub site [TODO:REF].

@@

\subsubsection{Productivity}
The theoretical maximum mean productivity ($\mu$M/h) was inversely proportional to the protein length ($l_{P}$) and varied between 1 and 12 $\mu$M/h for the proteins sampled (Fig.~\ref{fig:Prof}A-B). The theoretical maximum productivity was within a standard deviation with and without
amino acid supplementation for each protein, but varied significantly between proteins.
For instance, BMP10 (XX aa) had a optimal productivity of approximately 2.5 $\mu$M/h whereas CAT (ZZ aa) had an optimal productivity of approximately 12 $\mu$M/h.
To examine the influence of protein length further, the mean optimal productivity was plotted against the carbon number of each protein (Fig.~\ref{fig:Prof}B).
The optimal productivity and protein length were related by the power-law relationship $\alpha\times(l_{P})^{\beta}$, where $\alpha = 6.0\times 10^{6}$ [TODO: UNITS?] and $\beta=-1.9$.
Interestingly, CAT did not obey the power-law relationship; the relatively high productivity of CAT was due to its T7 promoter.
However, CAT expressed under a P70a promoter followed the power-law correlation with a productivity of approximately 8.8 $\mu$M/h (predicted to be XX by the optimal productivity correlation). Taken together, the maximum optimal productivity of a cell free reaction was found to be inversely proportional to protein size,
following a power-law relationship for proteins expressed under a P70a promoter.

@@

\subsubsection{Carbon Yield}
The theoretical maximum mean carbon yield was inversely proportional to protein length and varied between 40\% and 57\% for the proteins sampled (Fig.~\ref{fig:Prof}E-F).
The relationship between the optimal carbon yield and protein length was linear; $m_{Y}\times(l_{P})+b_{Y}$ where $m_{Y} =  -8.03\times10^{-3}$ [TODO: UNITS] with supplementation, and $m_{Y} = -6.98\times10^{-3}$ [UNITS] without. The y-intercept $b_{Y} = 63.83$ [UNITS] with supplementation, and $b_{Y} = 58.86$ [TODO:UNITS] without.
The linear carbon yield models showed good predictability for estimating the optimal carbon yield for a range of protein lengths ($R^{2} = 0.95$ with amino acids
and $R^{2} = 0.90$ without supplementation), irrespective of the promoter used to control protein expression.
The effective yield models were approximately parallel, with a drop in carbon yield of approximately 7\% without supplementation.
Only the necessary amount of amino acids were used for the production of the protein of interest; thus, it may be hypothesized that all the glucose was used to power CFPS and did not contribute to the carbon yield.
In that case, the carbon yield without the glucose contribution would be 100\%.
However, in the experimentally constrained case, CAT was produced with a carbon yield of 6.2\% compared to the theoretical maximum of 58\%.
This decrease in carbon yield suggests inefficiencies in CFPS that can potentially be improved.
Constraint based calculations assume a metabolic pseudo steady state; thus, intermediate metabolites cannot accumulate within the cell-free extract.
In addition, the flux balance analysis calculation is solved by maximizing the flux through the protein production reaction.
Therefore, carbon flux will travel through the network to optimize the maximum flux through the protein synthesis reaction.
In examining the experimental dataset, there is a high accumulation of organic acids, especially acetate, pyruvate and lactate.
The experimental performance could be improved by diverting this carbon toward the protein of interest by knockouts during the cell-free extract preparation.

@@

To investigate the connection between carbon yield and oxidative phosphorylation further,
we calculated the CAT carbon yield as a function of the oxidative phosphorylation flux (Fig.~\ref{fig:oxphos_yield}).
Oxidative phosphorylation had a strong effect on the carbon yield, both with and without amino acid supplementation.
In the presence of amino acid supplementation, the carbon yield ranged from 20\% to approximately 60\%,
depending upon the oxidative phosphorylation flux.
However, without amino acid supplementation, the carbon yield dropped to approximately 10\%, and reached a maximum of 50\%.
In the absence of supplementation, a lower carbon yield was expected for the same oxidative phosphorylation flux, as glucose was utilized for both energy generation and amino acid biosynthesis.
In all cases, whenever the carbon yield was below its theoretical maximum, there was an accumulation of both acetate and lactate.
The experimental dataset also exhibited a mixture of acetate and lactate accumulation during CAT synthesis, which suggested the CFPS reaction was not
operating with optimal oxidative phosphorylation activity.
Oxidative phosphorylation is a membrane associated process, however CFPS has no cell membrane.
Jewett et al hypothesized that membrane vesicles present in the CFPS reaction carry out oxidative phosphorylation [TODO: REF];
however, the number, size, protein loading and lifetime of these vesicles remains an open area of study.
Thus, we expect CFPS will have suboptimal oxidative phosphorylation capacity.

% The addition of phosphate showed an increase in CAT yield; however, it is unclear whether the addition of phosphate enhances oxidative phosphorylation, inhibits phosphatase reactions, or both \cite{Jewett:2008aa}.
% Interestingly, the addition of NADH did not increase the rate of protein synthesis, since the concentration of ATP was most likely saturated.
% An alternative strategy may be the inhibition of anaerobic processes in cell-free, in order to minimize unwanted byproducts such as acetate and lactate.
%Thus, it is interesting to see how much of the energetic needs of the system are met by oxidative phosphorylation and how much from anaerobic processes.
\begin{figure}[t!]
\includegraphics[width=0.7\textwidth]{./figs/Fig-4-OxyPhox-versus-Carbon-Yield.pdf}
\caption{The deGFP carbon yield versus oxidative phosphorylation flux, across an ensemble of 1000 flux balance solutions, for amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray).}
\label{fig:oxphos_yield}
\end{figure}

\subsection{Optimal Metabolic Flux Distribution}
Amino acid supplementation altered the optimal metabolic flux distribution predicted for CAT production (Fig.~\ref{fig:flux}).
To investigate the influence of amino acid supplementation, we compared the simulated metabolic flux distributions for CAT production with and without external amino acids.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-5-FluxDistribition-Optimal.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and oxidative phosphorylation, for three different cases: (A) amino acid uptake and synthesis, (B) amino acid uptake without synthesis, and (C) amino acid synthesis without uptake. Mean flux across the ensemble, normalized to glucose uptake flux. Thick arrows indicate flux to or from amino acids.}
\label{fig:flux}
\end{figure}
In the presence of amino acid supplementation, and \textit{de~novo} amino acid synthesis, there was an incomplete TCA cycle, where a combination
of glucose and amino acids powered protein expression (Fig.~\ref{fig:flux}A).
Glucose was consumed to produce acetyl-coenzyme A, and associated by-products,
while glutamate was converted to alpha-ketoglutarate which traveled to oxaloacetic acid and pyruvate for additional amino acid biosynthesis.
On the other hand, in the presence of amino acid supplementation, but without \textit{de~novo} amino acid biosynthesis, there was no TCA cycle flux.
In this case, ATP was produced by a combination of substrate level and oxidative phosphorylation, where
ubiquinone was regenerated via \textit{nuo} activity, without relying on succinate dehydrogenase in the TCA cycle (Fig.~\ref{fig:flux}B).
These first two cases where amino acids were available had similar performance, and their respective metabolic flux distributions had a 99\% correlation for all proteins. However, in the absence of amino acid supplementation (where all amino acids were synthesized \textit{de~novo} from glucose),
the energy efficiency and carbon yield decreased; in this case the TCA cycle was largely complete and there was diversion of metabolic flux into the Entner-Doudoroff pathway to produce NADPH (Fig.~\ref{fig:flux}C). [TODO: We need a link sentence to the experimentally constrained case.]

@@

The case constrained by experimental measurements (Fig. \ref{fig:flux_exp}), had the highest correlation of 0.66 with the flux distribution from the case supplied with no amino acids and a correlation of 0.52 with the cases supplied with amino acids.
Thus there are some differences in the flux distribution compared to the optimum solutions which may provide some insight to improve CFPS performance.
Metabolic fluxes were constrained by experimental measurements (available in Supporting Information) where available for the first hour
which constrained the feasible solution space.
The central carbon organic acids showed good agreement with the data (Fig.~\ref{fig:flux_exp}B).
Only certain amino acid synthesis reactions were blocked since during the growth of \emph{E.~coli} not all amino acids were supplied (see Materials and Methods).
During the cell-free reaction all amino acids were supplied, however glucose still traveled through all the major pathways, and the same metabolic precursors were still utilized for amino acid biosynthesis.
In this case, it is unclear which substrate (glucose or amino acids) is used to power CFPS and may in fact be a combination of both.
The optimum solutions only produced the required amount of amino acids necessary, however in examining the measurements there is an accumulation of alanine and glutamine which may explain some of the differences in the correlations of the flux distributions.
Accumulation of pyruvate, lactate, acetate, and other organic acids can be seen (Fig.~\ref{fig:flux_exp}B) and shows an inefficiency of carbon utilization.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-6-FluxDistribition-Experimental.pdf}
\caption{Constraint based simulation of CAT production for an experimentally constrained case. (A) Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and oxidative phosphorylation.  Mean flux across the ensemble, normalized to glucose uptake flux. Thick arrows indicate flux to or from amino acids. (B) Central carbon metabolite measurements versus simulations over a one hour time course.}
\label{fig:flux_exp}
\end{figure}

\subsubsection{Potential Alternative Metabolic Optima}
One of the interesting features of constraint based approaches is the ability to compute alternative optimal solutions.
These solutions have the same objective value, e.g., the productivity but different underlying metabolic flux distributions.
Despite the constraints by experimental measurements, it is difficult to calculate the physiological flux distribution of metabolism (Fig.~\ref{fig:norm})
For example, there is a high flux through the Entner-Douodoroff pathway, but this may be an artifact of the optimal solution.
A knockout of the Entner-Doudoroff pathway has no effect on the norm productivity of CAT compared to no knockouts (Fig.~\ref{fig:norm}A).
In addition, pairwise knockouts of Entner-Doudoroff and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) in the network result in the same optimal solution of CAT productivity.
However, there is a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B);
the flux balance approach will reroute the flux to optimize the objective function.
Interestingly, there are crucial subgroups that are detrimental to CAT productivity.
For example, a single group knockout of glycolysis/gluconoeogenesis or oxidative phosphorylation leads to no CAT production.
There were also cases of group knockouts that had no effect on productivity or its flux distribution such as isoleucine, leucine, and valine biosynthesis knocked out with histidine biosynthesis.
Flux balance analysis has been shown to have multiple alternative optimal solutions with flux variablitiy analysis and mixed-integer linear programming resulting in a poor depiction of the physiological flux distribution \cite{LEE2000711, Mahadevan2003264, Schuetz119}.
In our case, the constraint based simulation reached the same optimal solution of CAT productivity for 40\% of the pairwise group knockouts, but had a different flux distribution than compared to no knockouts.
To determine which reactions occur in CFPS, adding thermodynamic feasibility constraints to reactions may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013}.
It would also be interesting to track the carbon flux using C\textsuperscript{13} labeling in CFPS to better constrain the flux distribution.
However, while C\textsuperscript{13} labeling techniques are well established for \emph{in~vivo} processes \cite{Zamboni:2009},
they are an active area of research for cell free systems.
\begin{figure}[t!]
\includegraphics[width=1.0\textwidth]{./figs/Fig-7-FluxDistribition-Analysis.pdf}
\caption{The norm effect of pairwise knockouts of subgroups in the cell-free network (A) CAT productivity norm compared to no knockouts in the experimentally constrained case. (B) Flux distribution norm compared to no knockouts in the experimentally constrained case.}
\label{fig:norm}
\end{figure}

@@

\subsection*{Pairwise group knockouts.}
Pairwise and single group knockouts were simulated by setting the flux bounds for the all the reactions in a group to zero.
We grouped reactions in the cell-free network into 19 subgroups (available in Supporting Information).
We computed the norm of the productivity of CAT for each pairwise knockout compared to the productivity of CAT with no knockouts.
We also computed the norm of the flux distribution for each pairwise knockout compared to the flux distribution with no knockouts.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
